Numerical analysis of the linear vertical shear instability in astrophysical disks

### 2014-12-29 setup.tex
added beta cooling requirements
replace numerical section with that of full equations
restructing analytic discussion

The equilibrium disk satisfies
\begin{align}
  0 &= \frac{1}{\rho}\frac{\p P}{\p z} + \frac{\p \Phi_*}{\p z},\label{vert_equilibrium}\\
  r\Omega^2&= \frac{1}{\rho}\frac{\p P}{\p r} + \frac{\p \Phi_*}{\p
    r},\label{hori_equilibrium} 
\end{align} 
where $\Phi_* = -GM_*(r^2 + z^2)^{-1/2}$ is the gravitational
potential of the central star. Note that the disk is in general
baroclinic with $\Omega = \Omega(r,z)$, since
Eq. \ref{vert_equilibrium}---\ref{hori_equilibrium} imply, with
Eq. \ref{eqm_press},  
\begin{align}
  r\frac{\p\Omega^2}{\p z} = - \rho^{\Gamma-1}\frac{\p\ln\rho}{\p z}\frac{dK}{dr}.
\end{align}

@@

\subsection{Vertically isothermal disks}
For the special case $\Gamma=1$ the disk is isothermal in the
vertical direction. The equilibrium density $\rho(r,z)$ and rotation
$\Omega(r,z)$ is then 
\begin{align}
  \frac{\rho(r,z)}{\rho_0(r)} &=
  \exp{\left[\frac{1}{\epsilon^2}\left(\frac{1}{\sqrt{1+z^2/r^2}}-1\right)\right]}\\    
  \frac{\Omega^2(r,z)}{\Omega_k^2(r)}& =1+ (p+q)\epsilon^2(r) + q\left(1 -
    \frac{1}{\sqrt{1+z^2/r^2}}\right),
\end{align}
which may be obtained from Eq. \ref{eqm_dens} and Eq. \ref{eqm_rot} by
taking the limit $\Gamma\to 1$. The vertically isothermal disk has no
surface, but we can define a characteristic scale-height 
\begin{align}
  \Hiso \equiv \epsilon r.% \quad\quad(\Gamma \equiv 1),   
\end{align} 
and in this case the characteristic aspect-ratio $h = \epsilon$.

@@

We can eliminate the perturbed velocities between the linearized
equations to obtain a pair or ordinary differential equations
for $W,Q$ as
\begin{align}
  &\frac{\sigma^2}{c_s^2}Q + \frac{\sigma^2k_x^2}{D}W \notag\\ 
  &=\left(\frac{\ii
    k_x r}{D}\frac{d\Omega^2}{dz} -
  \frac{d\ln{\rho}}{dz}\right)\left[\frac{dW}{dz} +
  \frac{d\ln{\rho}}{dz}\left(W-Q\right)\right] \notag\\
&- \frac{d^2W}{dz^2} - \frac{d\ln{\rho}}{dz}\left(\frac{dW}{dz} -
  \frac{dQ}{dz}\right) - \frac{d^2\ln{\rho}}{dz^2}\left(W-Q\right),\label{ode_w}\\
&\sigma^2W - \frac{\gamma}{\Gamma}\sigma^2 Q +
\frac{\ii\sigma}{t_c}\left(W-\frac{Q}{\Gamma}\right)\notag\\
&=c_s^2\frac{d\ln{\rho}}{dz}\left(\frac{\gamma}{\Gamma} - 1\right) 
\left[\frac{dW}{dz} + \frac{d\ln{\rho}}{dz}\left(W-Q\right)\right],\label{ode_Q} 
\end{align}
 % It is convenient to express $\delta v_x$ as a function of $W$ and
 % $\delta v_z$,
 % \begin{align}
 %   \delta v_x = -\frac{1}{D}\left(\sigma k_x W +
 %     r\frac{d\Omega^2}{dz}\delta v_z\right),
 % \end{align}
where
\begin{align}
  D \equiv \kappa^2 - \sigma^2.
\end{align}
Eq. \ref{ode_w}---\ref{ode_Q} are convenient for numerical work, 
but alternative forms are more suitable for analytical discussion, and 
will be derived later.

\subsection{Boundary conditions}
In numerical calculations a finite vertical domain $z\in[-1,1]\zmax$
is adopted. For $\Gamma\neq1$ we take $\zmax$ to be slightly less than
$H$ to avoid the zero-density surface. At $z=\pm\zmax$ we impose
\begin{align}
  \delta v_z = 0,
\end{align}
for a solid boundary, or 
\begin{align}
  \Delta P \equiv \delta P + \bm{\xi}\cdot\nabla P= 0,  
\end{align}
for a free surface, where $\bm{\xi}$ is the Lagrangian
displacement.

>>>>

The equilibrium disk satisfies
\begin{align}
  0 &= \frac{1}{\rho}\frac{\p P}{\p z} + \frac{\p \Phi_*}{\p z},\label{vert_equilibrium}\\
  r\Omega^2&= \frac{1}{\rho}\frac{\p P}{\p r} + \frac{\p \Phi_*}{\p
    r},\label{hori_equilibrium} 
\end{align} 
where $\Phi_* = -GM_*(r^2 + z^2)^{-1/2}$ is the gravitational
potential of the central star. Note that the disk is in general
baroclinic with $\Omega = \Omega(r,z)$, since
Eq. \ref{vert_equilibrium}---\ref{hori_equilibrium} imply, with
Eq. \ref{eqm_press},  
\begin{align}
  r\frac{\p\Omega^2}{\p z} = - \rho^{\Gamma-1}\frac{\p\ln\rho}{\p z}\frac{dK}{dr}.\label{vertical_shear}
\end{align}

@@

\subsection{Vertically isothermal disks}
For the special case $\Gamma=1$ the disk is isothermal in the
vertical direction. The equilibrium density $\rho(r,z)$ and rotation
$\Omega(r,z)$ is then 
\begin{align}
  \frac{\rho(r,z)}{\rho_0(r)} &=
  \exp{\left[\frac{1}{\epsilon^2}\left(\frac{1}{\sqrt{1+z^2/r^2}}-1\right)\right]}\\    
  \frac{\Omega^2(r,z)}{\Omega_k^2(r)}& =1+ (p+q)\epsilon^2(r) + q\left(1 -
    \frac{1}{\sqrt{1+z^2/r^2}}\right),\label{vertiso_eqm}
\end{align}
which may be obtained from Eq. \ref{eqm_dens} and Eq. \ref{eqm_rot} by
taking the limit $\Gamma\to 1$. The vertically isothermal disk has no
surface, but we can define a characteristic scale-height 
\begin{align}
  \Hiso \equiv \epsilon r.% \quad\quad(\Gamma \equiv 1),   
\end{align} 
and in this case the characteristic aspect-ratio $h = \epsilon$.

@@

We can eliminate the perturbed velocities between the linearized
equations to obtain a pair or ordinary differential equations
for $W,Q$ as
\begin{align}
  &\frac{\sigma^2}{c_s^2}Q + \frac{\sigma^2k_x^2}{D}W \notag\\ 
  &=\left(\frac{\ii
    k_x r}{D}\frac{d\Omega^2}{dz} -
  \frac{d\ln{\rho}}{dz}\right)\left[\frac{dW}{dz} +
  \frac{d\ln{\rho}}{dz}\left(W-Q\right)\right] \notag\\
&- \frac{d^2W}{dz^2} - \frac{d\ln{\rho}}{dz}\left(\frac{dW}{dz} -
  \frac{dQ}{dz}\right) - \frac{d^2\ln{\rho}}{dz^2}\left(W-Q\right),\label{ode_w}\\
&\sigma^2W - \frac{\gamma}{\Gamma}\sigma^2 Q +
\frac{\ii\sigma}{t_c}\left(W-\frac{Q}{\Gamma}\right)\notag\\
&=c_s^2\frac{d\ln{\rho}}{dz}\left(\frac{\gamma}{\Gamma} - 1\right) 
\left[\frac{dW}{dz} + \frac{d\ln{\rho}}{dz}\left(W-Q\right)\right],\label{ode_Q} 
\end{align}
 % It is convenient to express $\delta v_x$ as a function of $W$ and
 % $\delta v_z$,
 % \begin{align}
 %   \delta v_x = -\frac{1}{D}\left(\sigma k_x W +
 %     r\frac{d\Omega^2}{dz}\delta v_z\right),
 % \end{align}
where
\begin{align}
  D \equiv \kappa^2 - \sigma^2.
\end{align}
% Eq. \ref{ode_w}---\ref{ode_Q} are convenient for numerical work, 
% but alternative forms are more suitable for analytical discussion, and 
% will be derived later.
